---
title: "BAF Data Discovery"
output: html_document
---

## Installations & Imports

```{r}
# Install Packages
#install.packages(pkgs = c("tidyverse","knitr","flextable","gtsummary","dlookr","emmeans","DHARMa", "purrr", "broom", "rcompanion", "ggeffects", "cardx", "officer"))

# Libraries & Imports
packages <- c("tidyverse","knitr","flextable","gtsummary","dlookr","emmeans", "DHARMa", "purrr", "broom", "gt", "rcompanion", "ggeffects","cardx", "officer")
invisible(lapply(X = packages, FUN = library, character.only = TRUE, quietly = TRUE))
```

## Reading Data

```{r}
# Load BAF Data
baf <- read.csv("Base.csv")
```

```{r}
# Create a subset n= 100,000
fraud_1 <- baf %>% filter(fraud_bool == 1)
fraud_0 <- baf %>% filter(fraud_bool == 0)

# Sample from each group
sample_1 <- fraud_1 %>% sample_n(2000)
sample_0 <- fraud_0 %>% sample_n(198000)

# Combine records
baf_subset <- bind_rows(sample_1, sample_0)
baf_subset <- baf_subset %>% slice_sample(n = nrow(baf_subset))
```

## Chi-squared Test of Independance

```{r}
cat_vars <- c("payment_type", "housing_status", "employment_status", "device_os")

# Function that performs the Chi-squared test for each categorical variable
chi_results <- map_dfr(cat_vars, function(var) {
  tbl <- table(baf_subset[[var]], baf_subset$fraud_bool)
  test <- chisq.test(tbl)
  cramers_v <- cramerV(tbl, bias.correct = FALSE)  
  
  formatted_p <- if (test$p.value < 1e-4) {"< 0.0001"} else {
    formatC(test$p.value, format = "f", digits = 4)}
  
  tibble(variable = var,
    chi_squared = round(test$statistic, 2),
    p_value = formatted_p,
    cramers_v = round(cramers_v, 3))
})

```

```{r}
# Display Results
chi_vars <- c("variable", "chi_squared", "p_value", "cramers_v")

flextable(chi_results) %>%
  set_caption("Chi-squared Test of Independance") %>%
  add_footer_lines("Each test was performed using the same 100,000 record sample.") %>%
  align(j = chi_vars, align = "center", part = "body") %>%
  align(part = "header", align = "center") %>%
  border_inner_h(part = "body", border = fp_border(color = "gray", width = 0.5)) %>%
  autofit()

  
```

The results above suggest there **is** a statistically significant association between each categorical variable and fraud. **However**, most variables display a very weak association with fraud when examining Cramers v (< 0.1).

---

## Fraud Rate per Category

```{r}
# Examining Fraud Rate for payment_type
table_data <- table(baf$payment_type, baf$fraud_bool)
round(prop.table(table_data, margin = 1), 3)
```

```{r}
# Examining Fraud Rate for housing_status
table_data <- table(baf$housing_status, baf$fraud_bool)
round(prop.table(table_data, margin = 1), 3)
```

```{r}
# Examining Fraud Rate for employment_status
table_data <- table(baf$employment_status, baf$fraud_bool)
round(prop.table(table_data, margin = 1), 3)
```

```{r}
# Examining Fraud Rate for device_os
table_data <- table(baf$device_os, baf$fraud_bool)
round(prop.table(table_data, margin = 1), 3)
```

The fraud rates don't show a pattern of high fraud for any particular category among all variables. Most categories have fraud rates of ~1% which is likely due to the overall small proportion of fraudulent records in the data.

---

## Univariate Analysis

```{r}
# Logistic Regression: Payment Type
m <- glm(fraud_bool ~ payment_type , family= binomial, data= baf)
```

```{r}
# Creating an informative model summary table

coefs <- round(summary(m)$coefficients, 3)
coef_df <- as.data.frame(coefs)
coef_df$Variable <- rownames(coef_df)
rownames(coef_df) <- NULL
colnames(coef_df)[1:4] <- c("Estimate", "Std_Error", "z_value", "p_value")

# Calculate odds ratio and predicted probability
coef_df$Odds_Ratio <- round(exp(coef_df$Estimate), 3)
coef_df$Predicted_Probability <- round(coef_df$Odds_Ratio / (1 + coef_df$Odds_Ratio), 3)

# Adjust variable names to reflect payment_type levels
coef_df$Variable <- gsub("payment_type", "", coef_df$Variable)
coef_df$Variable[coef_df$Variable == "(Intercept)"] <- "Intercept (AA)"

# Filter desired columns
summary_table1 <- coef_df[, c("Variable", "Estimate", "p_value", "Odds_Ratio", "Predicted_Probability")]

# Format p-values
summary_table1$p_value <- ifelse(summary_table1$p_value < 0.0001, "< 0.0001",
                                formatC(summary_table1$p_value, digits = 4, format = "f"))
# Future formatting
num_cols <- c("Variable","Estimate","p_value","Odds_Ratio","Predicted_Probability")
```

```{r}
flextable(summary_table1) %>%
  set_header_labels(Variable = "payment type", Estimate = "log odds", p_value = "p-value", Odds_Ratio = "odds ratio",
    Predicted_Probability = "predicted probability") %>%
  
    colformat_num(j = c("Estimate", "Odds_Ratio", "Predicted_Probability"), digits = 3) %>%
    set_caption("Payment Type Univariate Analysis Results") %>%
    add_footer_lines("Note: Payment type AA is the reference category for the log odds & odds ratios.") %>%
    font(fontname = "Arial", part = "header") %>%
  
    align(j = num_cols, align = "center", part = "body") %>%
    align(part = "header", align = "center") %>%
    border_inner_h(part = "body", border = fp_border(color = "gray", width = 0.5)) %>%
    autofit()
```

The model's results show that: 

- AC is 3.2 times or 420% more likely to be fraud compared to AA
- AB is 2.14 times or 314% more likely to be fraud compared to AA
- AD is 2 times or 306% more likely to be fraud compared to AA
- AE is 34% less likely to be fraud compared to AA

The final ordering of payment categories from **most to least** associated with fraud is **[AC, AB, AD, AE, AA]**. This ordering will allow us to interpret and verify our future classifier's results. 

---

```{r}
# Logistic Regression: Housing Status
m2 <- glm(fraud_bool ~ housing_status , family= binomial, data= baf)
```

```{r}
# Creating an informative model summary table

coefs <- round(summary(m2)$coefficients, 3)
coef_df <- as.data.frame(coefs)
coef_df$Variable <- rownames(coef_df)
rownames(coef_df) <- NULL
colnames(coef_df)[1:4] <- c("Estimate", "Std_Error", "z_value", "p_value")

# Add Odds Ratio and Predicted Probability
coef_df$Odds_Ratio <- round(exp(coef_df$Estimate), 3)
coef_df$Predicted_Probability <- round(coef_df$Odds_Ratio / (1 + coef_df$Odds_Ratio), 3)

# Adjust variable names to reflect housing_status levels
coef_df$Variable <- gsub("housing_status", "", coef_df$Variable)
coef_df$Variable[coef_df$Variable == "(Intercept)"] <- "Intercept (BA)"

# Keep only desired columns
summary_table2 <- coef_df[, c("Variable", "Estimate", "p_value", "Odds_Ratio", "Predicted_Probability")]

# Format p-values
summary_table2$p_value <- ifelse(summary_table2$p_value < 0.0001, "< 0.0001",
                                formatC(summary_table2$p_value, digits = 4, format = "f"))
```

```{r}
flextable(summary_table2) %>%
  set_header_labels(Variable = "housing status", Estimate = "log odds", p_value = "p-value", Odds_Ratio = "odds ratio", Predicted_Probability = "predicted probability") %>%
  
  colformat_num(j = c("Estimate", "Odds_Ratio", "Predicted_Probability"), digits = 3) %>%
  set_caption("Housing Status Univariate Analysis Results") %>%
  add_footer_lines("Note: Housing status BA is the reference category for the log odds & odds ratios.") %>%
  font(fontname = "Arial", part = "all") %>%
  
  align(j = num_cols, align = "center", part = "body") %>%
  align(part = "header", align = "center") %>%
  border_inner_h(part = "body", border = fp_border(color = "gray", width = 0.5)) %>%
  autofit()
```

The model's results show that:

- BB housing is 84.5% less likely to be fraud compared to BA
- BC housing is 84.1% less likely to be fraud compared to BA
- BD housing is 77.6% less likely to be fraud compared to BA
- BE housing is 91.1% less likely to be fraud compared to BA
- BF housing is 89.2% less likely to be fraud compared to BA
- BG housing is 89.8% less likely to be fraud compared to BA

The final ordering of housing categories from **most to least** associated with fraud is **[BA, BD, BC, BB, BF, BG, BE]**. This ordering will allow us to interpret and verify our future classifier's results. 

---

```{r}
# Logistic Regression: Employment Status
m3 <- glm(fraud_bool ~ employment_status , family= binomial, data= baf)
```

```{r}
# Creating an informative model summary table

coefs <- round(summary(m3)$coefficients, 3)
coef_df <- as.data.frame(coefs)
coef_df$Variable <- rownames(coef_df)
rownames(coef_df) <- NULL
colnames(coef_df)[1:4] <- c("Estimate", "Std_Error", "z_value", "p_value")

# Add Odds Ratio and Predicted Probability
coef_df$Odds_Ratio <- round(exp(coef_df$Estimate), 3)
coef_df$Predicted_Probability <- round(coef_df$Odds_Ratio / (1 + coef_df$Odds_Ratio), 3)

# Adjust variable names to reflect employment_status levels
coef_df$Variable <- gsub("employment_status", "", coef_df$Variable)
coef_df$Variable[coef_df$Variable == "(Intercept)"] <- "Intercept (CA)"

# Keep only desired columns
summary_table3 <- coef_df[, c("Variable", "Estimate", "p_value", "Odds_Ratio", "Predicted_Probability")]

# Format p-values
summary_table3$p_value <- ifelse(summary_table3$p_value < 0.0001, "< 0.0001",
                                formatC(summary_table3$p_value, digits = 4, format = "f"))
```

```{r}
flextable(summary_table3) %>%
  set_header_labels(Variable = "employment status", Estimate = "log odds", p_value = "p-value", Odds_Ratio = "odds ratio",Predicted_Probability = "predicted probability") %>%
  
  colformat_num(j = c("Estimate", "Odds_Ratio", "Predicted_Probability"), digits = 3) %>%
  set_caption("Employment Status Univariate Analysis Results") %>%
  add_footer_lines("Note: Employemnt status CA is the reference category for the log odds & odds ratios.") %>%
  font(fontname = "Arial", part = "all") %>%
  
  align(j = num_cols, align = "center", part = "body") %>%
  align(part = "header", align = "center") %>%
  border_inner_h(part = "body", border = fp_border(color = "gray", width = 0.5)) %>%
  autofit()
```

The model's results show that:

- CC is 2.2 times or 320% more likely to be fraud compared to CA
- CG is 1.1 times or 212% more likely to be fraud compared to CA
- CE is 84% less likely to be fraud compared to CA
- CF is 83 less likely to be fraud compared to CA
- CB is 44% less likely to be fraud compared to CA
- CD is 58.8% less likely to be fraud compared to CA


The final ordering of employment categories from **most to least** associated with fraud is **[CC, CG, CA, CB, CD, CF, CE]**. This ordering will allow us to interpret and verify our future classifier's results. 

---

```{r}
# Logistic Regression: Employment Status
m4 <- glm(fraud_bool ~ device_os , family= binomial, data= baf)
```

```{r}
# Creating an informative model summary table

coefs <- round(summary(m4)$coefficients, 3)
coef_df <- as.data.frame(coefs)
coef_df$Variable <- rownames(coef_df)
rownames(coef_df) <- NULL
colnames(coef_df)[1:4] <- c("Estimate", "Std_Error", "z_value", "p_value")

# Add Odds Ratio and Predicted Probability
coef_df$Odds_Ratio <- round(exp(coef_df$Estimate), 3)
coef_df$Predicted_Probability <- round(coef_df$Odds_Ratio / (1 + coef_df$Odds_Ratio), 3)

# Adjust variable names to reflect device os levels
coef_df$Variable <- gsub("device_os", "", coef_df$Variable)
coef_df$Variable[coef_df$Variable == "(Intercept)"] <- "Intercept (Linux)"

# Keep only desired columns
summary_table4 <- coef_df[, c("Variable", "Estimate", "p_value", "Odds_Ratio", "Predicted_Probability")]

# Format p-values
summary_table4$p_value <- ifelse(summary_table4$p_value < 0.0001, "< 0.0001",
                                formatC(summary_table4$p_value, digits = 4, format = "f"))
```


```{r}
flextable(summary_table4) %>%
  set_header_labels(Variable = "device os", Estimate = "log odds", p_value = "p-value", Odds_Ratio = "odds ratio",Predicted_Probability = "predicted probability") %>%
  
  colformat_num(j = c("Estimate", "Odds_Ratio", "Predicted_Probability"), digits = 3) %>%
  set_caption("Device Operating System Univariate Analysis Results") %>%
  add_footer_lines("Note: Device OS Linux is the reference category for the log odds & odds ratios.") %>%
  font(fontname = "Arial", part = "all") %>%
  
  align(j = num_cols, align = "center", part = "body") %>%
  align(part = "header", align = "center") %>%
  border_inner_h(part = "body", border = fp_border(color = "gray", width = 0.5)) %>%
  autofit()
```

The model's results show that:

- Windows os is 4.9 times or 589% more likely to be fraud compared to Linux
- Mac os is 2.7 times or 374% more likely to be fraud compared to Linux
- X11 os is 2.2 times or 318% more likely to be fraud compared to Linux
- Other os's are times or 212% more likely to be fraud compared to Linux


The final ordering of device os categories from **most to least** associated with fraud is **[Windows, Macintosh, X11, Other, Linux]**. This ordering will allow us to interpret and verify our future classifier's results. 

---

